---
title: 'STAT 140: Homework 4'
author: "YOUR NAME HERE"
date: "2/21/2020"
output:
  pdf_document: default
---

## BOOK EXERCISES (OpenIntro Statistics, Fourth Edition)

#### 3.5) **Coin flips.** For (a)-(c), define events (e.g. heads=H) and use appropriate probability notation and find the numerical answer. If you flip a fair coin 10 times, what is the probability of

a) getting all tails?

**Answer:**

b) getting all heads?

**Answer:**

c) getting at least one tails?

**Answer:**

d) What rules did you use to calculate the probabilities in (a)-(c)?

**Answer:**

#### 3.9) **Disjoint vs. independent.** In parts (a) and (b), identify whether the events are disjoint, independent, or neither (events cannot be both disjoint and independent).

a) You and a randomly selected student from your class both earn A's in this course.

**Answer:**

b) You and your class study partner both earn A's in this course.

**Answer:**

c) If two events can occur at the same time, must they be dependent? **Explain in 1 sentence.**

**Answer:**

#### 3.12) **School absences.** Data collected at elementary schools in DeKalb County, GA suggest that each year roughly 25\% of students miss exactly one day of school, 15\% miss 2 days, and 28\% miss 3 or more days due to sickness. 

a) What is the probability that a student chosen at random doesn't miss any days of school due to sickness this year?

**Answer:**

b) What is the probability that a student chosen at random misses no more than one day?

**Answer:**

c) What is the probability that a student chosen at random misses at least one day?

**Answer:**

d) If a parent has two kids at a DeKalb County elementary school, what is the probability that neither kid will miss any school? Note in one sentence any assumption you must make to answer this question.

**Answer:**

e) If a parent has two kids at a DeKalb County elementary school,what is the probability that both kids will miss some school, i.e. at least one day? Note in one sentence any assumption you make. 

**Answer:**

f) If you made an assumption in part (d) or (e), do you think it is reasonable? If you didn't make any assumptions, double check your earlier answers.

**Answer:**

#### 3.13) **Joint and conditional probabilities.** $P(A)=0.3$, $P(B)=0.7$

a) Can you compute $P(A \text{ and } B)$ if you only know $P(A)$ and $P(B)$?

**Answer:**

b) Assuming that events $A$ and $B$ arise from an independent random process,

------

  i) what is $P(A \text{ and } B)$?

**Answer:**

  ii) what is $P(A\text{ or }B)$?

**Answer:**

  iii) what is $P(A|B)$?

**Answer:**

c) If we are given that $P(A \text{ and } B)=0.1$, are the random processes giving rise to events $A$ and $B$ independent?

**Answer:**

d) If we are given that $P(A \text{ and } B)=0.1$, what is $P(A|B)$?

**Answer:**

#### 3.39) Each row in the table on page 129 is a proposed grade distribution for a class. Identify each as a valid or invalid probability distribution, and explain your reasoning.

a)

b)

c)

d)

e)

f)


## R EXERCISE

The Law of Large Numbers states that as the sample size grows (as more observations are recorded), the observed probability of an event, $\hat{p}_n$, converges to the probability of that event, $p$. 

In class, we looked at this phenomenon for a fair coin. Here we will look at this phenomenon for a fair die.

*Sample function:*

The sample function allows us to simulate a random process, so we can ``flip coins" or ``roll dice" on the computer instead of in real life. Looking at the help file (?base::sample), you see that sample() has four arguments:

- **x**: for us this will be a vector of possible outcomes of our random variable
- **size**: n, the sample size (number of observations) - just an integer
- **replace**: a logical; we will set replace=TRUE so that on each die roll we can observe any one of the six outcomes
- **prob**: we do not need to specify this unless some observations are more likely than others. The default is to give equal weight/chance to all possible outcomes. 

```{r sample function}
## Demonstrate sample function for fair dice rolls

## Specify set of all outcomes in a vector
roll_outcomes <- c(1,2,3,4,5,6)

## Roll a fair die once
r1 <- sample(x=roll_outcomes, size=1, replace=TRUE)

r1

## Roll a fair die once
r5 <- sample(x=roll_outcomes, size=5, replace=TRUE)

r5
```
Describe what r1 gives you.

**Answer:**

If you roll again (rerun the sample line for r1), are you guaranteed get the same result? (Try it if you are not sure!) Why or why not? (**Limit answer to 1 sentence.**)

**Answer:**

Describe what r5 gives you.

**Answer:**

*Sample function in a loop:*

It is tedious to write a separate line of code each time I want to change my sample size. In R, we can write a loop that lets us increase the sample size automatically. I will not ask you to do this independently, but you should be aware that you can.

Here we are only going to keep track of the probability that we roll a 1 and plot the result.
```{r LLN fair die roll, fig.cap="Fair die rolls", out.width = '70%', fig.align='center', tidy=TRUE}
## This is the maximum number of observations I am going to make.
N <- 1000

## This creates a place to store my observations
## It is a data frame with two columns, n and p_n
## n is the sample size
## p_n is the observed proportion of times we roll a 1.
LLN_dr <- data.frame(n=1:N, p_n=rep(0,N))
for (n in 1:N){
  LLN_dr$p_n[n] <- sum(sample(x=roll_outcomes,size=n,replace=TRUE)==1)/n
}

## load ggplot2
library(ggplot2)

## ggplot for observed probability versus number of rolls
lln_p <- (ggplot(data=LLN_dr, aes(x=n, y=p_n)) + geom_line()
          + theme_bw() 
          + geom_hline(yintercept = 1/6, linetype="dashed", color="red")
          ## Adds a red dashed line at p, the true probability
          + xlab("n (number of rolls)") + ylab(expression(hat(p)[n])))
lln_p
```

Now, think about a weighted die (this means that each of the six numbers on the six faces of the die does not have the same chance of appearing). Simulate the process (use the sample function - you can copy and paste the code from the LLN fair die roll R chunk and then tweak it as necessary for our new problem). Keep track of the probability that we roll a 1 and plot the result as we did before. You will need to add prob=weighted.prob within the parentheses for sample, where weighted.prob is specified below: 

**Answer:**

```{r LLN weighted die roll, fig.cap="Weighted die rolls", out.width = '70%', fig.align='center', tidy=TRUE}
## Weighted probability vector (Slide 63)
weighted.prob <- c(3/6,1/6,1/12,1/12,1/12,1/12)

## Copy and paste code from previous R chunk here
## Make sure you add the fourth argument, prob=weighted.prob, to the sample function
## Make sure you change the value of yintercept within geom_hline to correspond to the expected probability of observing a 1.

## -------------- YOUR ANSWER HERE -------------- ##


```



